Lessons Learned about JWST Transiting Planet Time Series from Ground-based Studies of the NIRCam Detectors

### 2020-04-29 ms.tex
updating covariance weighting method and results sections

\subsubsection{Empirical Covariance Matrix}\label{sec:empCovarianceMatrix}
The ideal weighting of the spectrum from Equation \ref{eq:covarianceweights} depends on the covariance matrix $C_{ij}$.
We use CDS (correlated double sampling) read pairs of OTIS long dark data (108 frames) with reference pixel corrections to estimate $C_{ij}$.
We consider a wavelength bin that is 100 pixels in the dispersion direction (0.1$\mu$m) and 14 spatial spatial pixels.
Figure \ref{fig:covSpecSpatial} shows the covariance matrix of this wavelength bin, taken from the median of 20 regions of the detector.
The diagonal elements are consistent with published NIRCam detector performance tables with $\sigma^2 \approx 13^2 (e^-)^2=169 (e^-)^2$.
It is clear that 1/f noise causes covariance most prominently along the spectral direction (the fast-read direction), with a covariance nearlynear 36~$(e^-)^2$.
Surprisingly, there is also a covariance between spatial pixels (the slow read direction) of $\sim$20 $(e^-)^2$.

\begin{figure}[!hbtp]
\centering
\includegraphics[width=.6\columnwidth]{spec_spatial_cov.pdf}
\caption{The read noise shows significant correlations between pixels (unlike the traditional assumption that each pixel's noise is statistically independent).
The data was computed from a long dark exposure of 54 read pairs where reference pixel corrections and column-by-column and row-by-row subtractions have been performed.
The detector's diagonal elements are consistent with expected performance of $\approx 13^2 (e^-)^2=169 (e^-)^2$.
Spatially, the pixels' have a strong covariance of 36~$(e^-)^2$ and a somewhat smaller spatial covariance of 20$(e^-)^2$.}\label{fig:covSpecSpatial}
\end{figure}

@@

The empirical covariance matrix from Figure \ref{fig:covSpecSpatial} is used to estimate the optimal pixel weights using Equation \ref{eq:covarianceweights}.
We reduce the noise and jaggedness of the empirical covariance matrix in Figure \ref{fig:covSpecSpatial} by averaging the value along the diagonal, across all spectral pixels and across the spatial pixels with the same value of $i -j$.
The resulting optimal weighting function $w_j/P_j$ is plotted in Figure \ref{fig:covWeightedSummation}.
Here, the profile is the value from \texttt{pynrc}, which uses \texttt{webbpsf}.
This optimal weighting significantly increases the contributions from the inner pixels

@@

>>>>

\subsubsection{Empirical Covariance Matrix}\label{sec:empCovarianceMatrix}
The ideal weighting of the spectrum from Equation \ref{eq:covarianceweights} depends on the covariance matrix $C_{ij}$.
We use CDS (correlated double sampling) read pairs of OTIS long dark data (108 frames) with reference pixel corrections to estimate $C_{ij}$.
We consider a wavelength bin that is 100 pixels in the dispersion direction (0.1$\mu$m) and 14 spatial spatial pixels.
We also subtract the CDS frames by the median of each row-by-row and the median of each column column-by-column to reduce the amplifier offsets and 1/f noise as done in Figure \ref{fig:longDarkPhot}.

Figure \ref{fig:covSpecSpatial} shows the covariance matrix of this wavelength bin, taken from the median covariance matrix of 20 regions of the detector.
The diagonal elements $\sim 150 (e^-)^2$ are smaller than published NIRCam detector performance tables ($\sigma^2 \approx 13^2 (e^-)^2=169 (e^-)^2$).
However, if we skip the row-by-row and column-by-column subtraction step, they are consistent with $\sigma^2 \approx 13^2 (e^-)^2=169 (e^-)^2$.
It is clear that 1/f noise causes covariance most prominently along the spectral direction (the fast-read direction), with a covariance near 20~$(e^-)^2$.
Surprisingly, there is also a covariance between spatial pixels (the slow read direction) of $\sim$10 $(e^-)^2$.

\begin{figure}[!hbtp]
\centering
\includegraphics[width=.6\columnwidth]{spec_spatial_cov.pdf}
\caption{The read noise shows significant correlations between pixels (unlike the traditional assumption that each pixel's noise is statistically independent).
The data was computed from a long dark exposure of 54 read pairs where reference pixel corrections and column-by-column and row-by-row subtractions have been performed.
The detector's diagonal elements are better than the read noise squared due to the row-by-row and column-by-column subtraction step.
Spatially, the pixels' have a covariance of 20~$(e^-)^2$ and a somewhat smaller spatial covariance of 10$(e^-)^2$.}\label{fig:covSpecSpatial}
\end{figure}

@@

The empirical covariance matrix from Figure \ref{fig:covSpecSpatial} is used to estimate the optimal pixel weights using Equation \ref{eq:covarianceweights}.
We reduce the noise and jaggedness of the empirical covariance matrix in Figure \ref{fig:covSpecSpatial} by averaging the value along the diagonal, across all spectral pixels and across the spatial pixels with the same value of $i -j$.
The resulting optimal weighting function $w_j/P_j$ is plotted in Figure \ref{fig:covWeightedSummation}.
Here, the profile is the value from \texttt{pynrc}, which uses \texttt{webbpsf}.
This optimal weighting significantly increases the contributions from the inner pixels compared to variance weights.
The optimal weighting scheme also negatively weights some pixels outside the core of the PSF to reduce correlated noise.

@@

Optimal covariance weighting significantly lowers the contribution of read noise to the overall error budget.
We test this with a series of grism images with no photon noise added.
We do background subtraction along the dispersion and spatial directions.
We perform sum extraction over a 15 spatial pixels and then combine together 170 spectral pixels into a 0.17 $\mu$m wavelength bin.
The standard deviation of sum extraction is 3,860 DN, which is 2.7$\times$ the read noise, as shown in Table \ref{tab:noiseSummary1overf}.
The standard deviation of the extracted flux drops to 3,060 DN if optimal extraction is used.

We use Equation \ref{eq:covarianceweights}, to further beat down the read noise.
For the covariance weights, we assume constant off-diagonal weights of 15 (e$^-)^2$ and do not include the spectral covariance.
The resulting standard deviation of 850 DN is now 60\% of the photon noise.
Thus, we recommend optimal covariance weighting when extracting grism time series.
As the number of groups of the ramp is increased, the read noise should drop as 1/$\sqrt{N_G}$ so the covariance weighting most important when $N_G$ is small.

The methodology in calculating $P_j$ and Cov($S_i,S_j$) can significantly affect the precision on the time series.
Typically $P_j$ and the photon noise contribution to Cov($S_i,S_j$) can be found by fitting smooth functions (polynomials) along the dispersion direction.
If $P_j$ is calculated from each image individually, it will inherit the 1/f noise variations of each image.
Therefore, we hold $P_j$ constant along the time series with a value saved from the middle image at all wavelengths.
Alternatively, $P_j$ could be calculated from a PSF library or theoretically from \texttt{WebbPSF} \citep{perrin2014webbpsf}.
